---
title: "p8105_hw2_hz2770"
author: "Hao Zheng"
date: "2021/10/9"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)

# get the path of the dataset.
getwd()
```

## Problem 1
```{r}
# Read and clean the Mr.Trash Wheel sheet
mr_trash_wheel_data = 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "Mr. Trash Wheel",
             range = cell_cols("A2:N534")) %>% 
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0))
```


```{r}
# Read and clean precipitation data for 2018 and 2019
precipitation_data_2018 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  drop_na(total) %>% 
  mutate(year = "2018")
```

```{r}
precipitation_data_2019=
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "2019 Precipitation",
    skip = 1
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  drop_na(total) %>% 
  mutate(year = "2019")
```


```{r}
# Then combine the two dataset
precipitation_data = 
  bind_rows(precipitation_data_2018, precipitation_data_2019) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year)
```

## 


## Problem 2

```{r}
# clean the data in pols-month.csv
pols_month = read_csv(
  "./fivethirtyeight_datasets/pols-month.csv"
) %>%
  separate(mon,c("year","month","day")) %>%
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day)) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(prez_gop = na_if(prez_gop,2)) %>% 
  drop_na(prez_gop) %>% 
  mutate(president = ifelse(prez_gop == 1, "gop", "dem")) %>% 
  select(-prez_gop, -prez_dem, -day)

pols_month
```

Second, clean the data in snp.csv using a similar process
```{r}
snp_data = read_csv(
  "./fivethirtyeight_datasets/snp.csv"
) %>%
  separate(date,c("month","day","year")) %>%
  mutate(year = as.integer(year) + 2000) %>% 
  mutate(year = ifelse(year >= "2020", year - 100, year)) %>% 
  mutate(month = as.integer(month)) %>%
  arrange(year,month) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month) %>% 
  select(-day)

snp_data
```

Now let's clean the unemployment data.
```{r}
unemployment_df = read_csv(
  "./fivethirtyeight_datasets/unemployment.csv"
)
unemployment_data = 
  pivot_longer(
    unemployment_df,
    Jan:Dec,
    names_to = "month",
    values_to = "perc_unemployment"
) %>% 
  janitor::clean_names() %>% 
  mutate(month = match(month,month.abb)) %>% 
  mutate(month = month.name[month])

unemployment_data
```

Join the dataset
```{r}
pols_snp_data = 
  left_join(pols_month, snp_data, by = c("year","month"))

pols_snp_unem_data = left_join(pols_month, unemployment_data, by = c("year","month"))
```


## Problem 3

First, clean the dataset.
```{r}
pop_baby_names = 
   read_csv("./data/Popular_Baby_Names.csv") %>% 
   janitor::clean_names() %>% 
# clean all the cases in dataset and moderate the ethnicity
   mutate(gender = tolower(gender),
          ethnicity = gsub(" ", "_", ethnicity),
          ethnicity = tolower(ethnicity),
          childs_first_name = str_to_title(childs_first_name)) %>% 
   mutate(ethnicity = recode(ethnicity, "asian_and_pacific_islander" = "asian_and_pacific", "white_non_hispanic" = "white_non_hisp", "black_non_hispanic" = "black_non_hisp")) %>%
#delete duplicate rows
  distinct()

pop_baby_names
```


Produce a table showing the rank in popularity of the name “Olivia” as a female baby name over time.
```{r}
Olivia_df = 
  pop_baby_names %>% 
  filter(childs_first_name == "Olivia") %>% 
  filter(gender == "female") %>% 
  select(year_of_birth, ethnicity, rank) %>% 
  arrange(year_of_birth,ethnicity)

Olivia_df
```


Produce a similar table showing the most popular name among male children.
```{r}
pop_male_names = 
  filter(pop_baby_names, gender == "male", rank == "1") %>%
  select(year_of_birth, ethnicity, childs_first_name) %>% 
  arrange(year_of_birth, ethnicity) %>% 
  rename(most_pop_male_name = childs_first_name)

pop_male_names
```

At last, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
scatter_df =  
  filter(pop_baby_names, year_of_birth == "2016", gender == "male", ethnicity == "white_non_hisp") 

df_plot = ggplot(scatter_df, aes(x = rank, y = count)) + 
  geom_point(color = 'red') + 
  labs(title = "Number of children with the names vs Rank of Popularity of Names for White Non-hispanic male Children Born in 2016",
       x = "rank",
       y = "number")

df_plot
```

